Weighted Model Counting with Conditional Measures

### 2020-07-21 paper.tex
Partial updates to the old proofs

\begin{itemize}
\item The Main Narrative
  \begin{enumerate}
  \item When weights are defined on literals, the measure on the free BA is
    fully independent.
  \item This means that the BA itself must be larger (i.e., have additional
    `meaningless' literals) to turn any probability distribution into an
    independent one.
  \item We show how we can define conditional weights on literals, allowing us
    to encode any probability distribution into a Boolean algebra that's not
    necessarily independent and thus can be smaller.
  \item We demonstrate a specific example of this by presenting a new way to
    encode Bayesian networks into instances of WMC and adapting a WMC algorithm
    (ADDMC) to run on the new format.
  \item We show that this results in significantly faster inference.
  \item We show that our encoding results in asymptotically fewer literals and
    fewer ADDs, and thus a simpler problem.
  \item (Maybe) we experimentally demonstrate a phase transition based on the
    number of variables per ADD.
  \end{enumerate}
\item Potential criticism may be that this is a step backwards and doesn't allow
  us to use SAT-based techniques for probabilistic inference. However, they can
  still be used for the 'theory+query' part.
  \begin{itemize}
  \item Zero-probability weights and one-probability weights can be interpreted
    as logical clauses. This doesn't affect ADDMC but could be useful for other
    solvers.
  \end{itemize}
\item[F] What are the main claims, what are the main takeaways, intuitive [???]
  of theorems to follow. To do this, we appeal to algebraic constructions to
  define the main concepts for introducing measures on Boolean algebras.
\item
  Algorithms\footnote{\url{http://beyondnp.org/pages/solvers/model-counters-exact/}}
  \begin{itemize}
  \item ADDMC \cite{DBLP:conf/aaai/DudekPV20} (rediscovered the
    multiplicativity of BAs in different words) (with optimal settings)
  \item Cachet \cite{DBLP:conf/sat/SangBBKP04}
  \item c2d \cite{DBLP:conf/ecai/Darwiche04}
  \item d4 \cite{DBLP:conf/ijcai/LagniezM17} (closed source, boo!)
  \item miniC2D  \cite{DBLP:conf/ijcai/OztokD15}
  \end{itemize}
\item Notable previous/related work
  \begin{itemize}
  \item Hailperin's approach to probability logic
    \cite{DBLP:journals/ndjfl/Hailperin84}
  \item Nilsson's (somewhat successful) probabilistic logic
    \cite{DBLP:journals/ai/Nilsson86,DBLP:journals/ai/Nilsson93}
  \item Logical induction: a big paper with a good overview of previous attempts
    to assign probabilities to logical sentences in a sensible way
    \cite{DBLP:journals/eccc/GarrabrantBCST16}
  \item Measures on Boolean algebras
    \begin{itemize}
    \item On possibility and probability measures in finite Boolean algebras
      \cite{DBLP:journals/soco/CastineiraCT02}
    \item Representation of conditional probability measures
      \cite{krauss1968representation}
    \end{itemize}
  \end{itemize}
\item Intuitively, a measure is just like a probability, except it's in
  $\mathbb{R}_{\ge 0}$ instead of $[0, 1]$.
\item ADDMC should probably be in textsc mode.
\end{itemize}

@@

\begin{itemize}
\item Make up my mind about $a, b$ vs. $x, y$ and stick to it (maybe $x, y$?).
\item Terminology: `with generating set $S$' $\to$ `over $S$'.
\item Notation: if $L$ denotes literals, then it doesn't denote a generating
  set. Literals should be $U$.
\end{itemize}

\begin{table}
  \caption{(in the same order)}
  \centering
  \begin{tabular}{lcc}
    \toprule
    & Set-theoretic notation & Boolean-algebraic notation \\
    \midrule
    Atoms (elements of $U$) & $a, b$ & $a, b$ \\
    \rowcolor{gray!10} Models (elements of $2^U$) & $\emptyset, \{a\}, \{b\}, \{a, b\}$ & $\neg a \land \neg b, a \land \neg b, \neg a \land b, a \land b$ \\
    & $\{ \emptyset, \{a\}, \{b\}, \{a, b\} \}$ & $\top$ \\
    & $\{ \emptyset, \{a\}, \{b\} \}, \{ \emptyset, \{a\}, \{a, b\} \}$ & $\neg a \lor \neg b, a \to b$ \\
    & $\{ \emptyset, \{b\}, \{a, b\} \}, \{ \{a\}, \{b\}, \{a, b\} \}$ & $b \to a, a \lor b$ \\
    & $\{\emptyset, \{a\}\}, \{\emptyset, \{b\}\}, \{\emptyset, \{a, b\}\}$ & $\neg b, \neg a, a \leftrightarrow b$ \\
    & $\{\{a\}, \{b\}\}, \{\{a\}, \{a, b\}\}, \{\{b\}, \{a, b\}\}$ & $(a \land \neg b) \lor (b \land \neg a), a, b$ \\
    & $\{\emptyset\}, \{\{a\}\}, \{\{b\}\}, \{\{a, b\}\}$ & $\neg a \land \neg b, a \land \neg b, \neg a \land b, a \land b$ \\
    \multirow{-7}{*}{Formulas (elements of $2^{2^U}$)} & $\emptyset$ & $\bot$ \\
    \bottomrule
  \end{tabular}
\end{table}

\begin{definition} \label{def:ba}
  A \emph{Boolean algebra} (BA) is a tuple $(\mathbf{B}, \land, \lor, \neg, 0,
  1)$ consisting of a set $\mathbf{B}$ with binary operations \emph{meet}
  $\land$ and \emph{join} $\lor$, unary operation $\neg$ and elements $0, 1 \in
  \mathbf{B}$ such that the following axioms hold for all $a, b, \in
  \mathbf{B}$:
  \begin{itemize}
  \item both $\land$ and $\lor$ are associative and commutative;
  \item $a \lor (a \land b) = a$, and $a \land (a \lor b) = a$;
  \item $0$ is the identity of $\lor$, and $1$ is the identity of $\land$;
  \item $\lor$ distributes over $\land$ and vice versa;
  \item $a \lor \neg a = 1$, and $a \land \neg a = 0$.
  \end{itemize}
\end{definition}

For clarity and succinctness, we will occasionally use three other operations
that can be defined using the original three\footnote{We use $+$ to denote
  symmetric difference because it is the additive operation of a Boolean ring.}:
\begin{align*}
  a \to b &= \neg a \lor b, \\
  a \leftrightarrow b &= (a \land b) \lor (\neg a \land \neg b), \\
  a + b &= (a \land \neg b) \lor (\neg a \land b).
\end{align*}
We can also define a partial order $\le$ on $\mathbf{B}$ as $a \le b$ if $a = b
\land a$ (or, equivalently, $a \lor b = b$) for all $a, b \in \mathbf{B}$.
Furthermore, let $a < b$ denote $a \le b$ and $a \ne b$. For the rest of this
paper, let $\mathbf{B}$ refer to the BA $(\mathbf{B}, \land, \lor, \neg, 0, 1)$.
For any $S \subseteq \mathbf{B}$, we write $\bigvee S$ for $\bigvee_{x \in S} x$
and call it the \emph{supremum} of $S$. Similarly, $\bigwedge S = \bigwedge_{x
  \in S} x$ is the \emph{infimum}. By convention, $\bigwedge \emptyset = 1$ and
$\bigvee \emptyset = 0$. For any $a, b \in \mathbf{B}$, we say that $a$ and $b$
are \emph{disjoint} if $a \land b = 0$.

\begin{definition}[\cite{DBLP:books/daglib/0090259,levasseur2012applied}]
  An element $a \ne 0$ of $\mathbf{B}$ is an \emph{atom} if, for all $x \in
  \mathbf{B}$, either $x \land a = a$ or $x \land a = 0$. Equivalently, $a \ne
  0$ is an atom if there is no $x \in \mathbf{B}$ such that $0 < x < a$. We say
  that $\mathbf{B}$ is \emph{atomic} if for every $a \in \mathbf{B} \setminus \{0
  \}$, there is an atom $x$ such that $x \le a$.
\end{definition}

\begin{lemma}[\cite{ganesh2006introduction}]
  For any two distinct atoms $a$, $b \in \mathbf{B}$, $a \land b = 0$.
\end{lemma}

@@

\begin{lemma}[\cite{givant2008introduction}] \label{lemma:atomic}
  All finite BAs are atomic.
\end{lemma}

@@

\begin{itemize}
\item[F2] Explain with examples: models $=$ elements [atoms] of algebra.
\item Be careful about mentioning ideals and quotients.
\item Note that we're going to use logical notation for BAs. Give an example of
  the two notations.
\end{itemize}

\begin{example} \label{example:construction}
  % TODO: keep this, but might need a lot of editing.
  Let $\mathcal{L}$ be a propositional logic with $p$ and $q$ as its only atoms.
  Then $L = \{ p, q, \neg p, \neg q \}$ is its set of literals. Let $w : L \to
  \mathbb{R}_{\ge 0}$ be the \emph{weight function} defined by
  \begin{align*}
    w(p) = 0.3, \\
    w(\neg p) = 0.7, \\
    w(q) = 0.2, \\
    w(\neg q) = 0.8.
  \end{align*}
  Let $\Delta$ be a theory in $\mathcal{L}$ with a sole axiom $p$. Then
  $\Delta$ has two models, i.e., $\{ p, q \}$ and $\{ p, \neg q \}$. The
  \emph{weighted model count} (WMC) \cite{DBLP:journals/ai/ChaviraD08} of $\Delta$ is
  then
  \[
    \sum_{\omega \models \Delta} \prod_{\omega \models l} w(l) =
    w(p)w(q) + w(p)w(\neg q) = 0.3.
  \]

@@

\begin{definition} \label{def:wmc}
  Let $U$ be an arbitrary set. Then
  \begin{itemize}
  \item a \emph{measure} is a function $M\colon 2^{2^U} \to \mathbb{R}_{\ge 0}$,
  \item a \emph{weight function} is a function $W\colon 2^U \to \mathbb{R}_{\ge
      0}$.
  \item a weight function is \emph{factored} if $W = \prod_{x \in U} W_x$ for
    some functions $W_x\colon 2^{\{x\}} \to \mathbb{R}_{\ge 0}$.
  \item every weight function induces a measure:
    \[
      M_W(x) = \begin{cases}
        0 & \text{if } x = \emptyset \\
        W(u) & \text{if } x = \{ u \} \\
        \sum_{\{u\} \le x} W(u) & \text{otherwise.}
      \end{cases}
    \]
    The process of calculating the value of $M_W(x)$ for some $x \in 2^{2^U}$
    with a given definition of $W$ is known as \emph{weighted model counting}.
  \end{itemize}
\end{definition}

@@

\section{Limitations of Factored Weighted Model Counting}

@@

\begin{lemma} \label{lemma:before_theorem}
  For any measure $m\colon \mathbf{B} \to \mathbb{R}_{\ge 0}$ and elements $a, b
  \in \mathbf{B}$,
  \begin{equation} \label{eq:to_prove}
    m(a \land b) = m(a)m(b)
  \end{equation}
  if and only if
  \begin{equation} \label{eq:to_prove2}
    m(a \land b) \cdot m(\neg a \land \neg b) = m(a \land \neg b)
    \cdot m(\neg a \land b).
  \end{equation}
\end{lemma}
\begin{proof}
  First, note that $a = (a \land b) \lor (a \land \neg b)$ and $(a \land b)
  \land (a \land \neg b) = 0$, so, by properties of a measure,
  \begin{equation} \label{eq:temp}
    m(a) = m(a \land b) + m(a \land \neg b).
  \end{equation}
  Applying \cref{eq:temp} and the equivalent expression for $m(b)$ allows us
  to rewrite \cref{eq:to_prove} as
  \[
    m(a \land b) = [m(a \land b) + m(a \land \neg b)][m(a \land b) + m(\neg a
    \land b)]
  \]
  which is equivalent to
  \begin{equation} \label{eq:temp6}
    m(a \land b)[1 - m(a \land b) - m(a \land \neg b) - m(\neg a \land b)] = m(a
    \land \neg b)m(\neg a \land b).
  \end{equation}
  Since $a \land b$, $a \land \neg b$, $\neg a \land b$, $\neg a \land \neg b$
  are pairwise disjoint and their supremum is $1$,
  \[
    m(a \land b) + m(a \land \neg b) + m(\neg a \land b) + m(\neg a \land \neg
    b) = 1,
  \]
  and this allows us to rewrite \cref{eq:temp6} into \cref{eq:to_prove2}. As all
  transformations are invertible, the two expressions are equivalent.
\end{proof}

\todo[inline]{This theorem needs a special case for zero weights.}
\begin{theorem}
  Let $\mathbf{B}$ be a free BA over $\{ l_i \}_{i=1}^n$ (for some $n \in
  \mathbb{N}$) with measure $m\colon \mathbf{B} \to \mathbb{R}_{\ge 0}$, and let
  $L = \{ l_i \}_{i = 1}^n \cup \{ \neg l_i \}_{i = 1}^n$. Then there exists a
  weight function $w\colon L \to \mathbb{R}_{\ge 0}$ such that $m = \WMC_w$ if
  and only if
  \begin{equation} \label{eq:wmccondition}
  m(l \land l') = m(l)m(l')
  \end{equation}
  for all distinct $l, l' \in L$ such that $l \ne \neg l'$.
\end{theorem}

\begin{remark}
  Note that if $n = 1$, then \cref{eq:wmccondition} is vacuously satisfied and
  so any valid measure can be expressed as WMC.
\end{remark}

\begin{proof}
  ($\Leftarrow$) Let $w\colon L \to \mathbb{R}_{\ge 0}$ be defined by
  \begin{equation} \label{eq:assumption}
    w(l) = m(l)
  \end{equation}
  for all $l \in L$. We are going to show that $\WMC_w = m$. First, note that
  $\WMC_w(0) = 0 = m(0)$ by the definitions of both $\WMC_w$ and $m$. Second,
  let
  \begin{equation} \label{eq:def_of_a}
    a = \bigwedge_{i=1}^n a_i
  \end{equation}
  be an atom in $\mathbf{B}$ such that $a_i \in \{ l_i, \neg l_i \}$ for all $i
  \in [n]$. Then
  \[
    \WMC(a) = \prod_{i=1}^n w(a_i) = \prod_{i=1}^n m(a_i) = m
    \left(\bigwedge_{i=1}^n a_i \right) = m(a)
  \]
  by \cref{def:wmc,eq:assumption,eq:wmccondition,eq:def_of_a}. Finally, note
  that if $\WMC$ and $m$ agree on all atoms, then they must also agree on all
  other non-zero elements of the Boolean algebra.

  ($\Rightarrow$) For the other direction, we are given a weight function
  $w\colon L \to \mathbb{R}_{\ge 0}$ that induces a measure $m = \WMC_w\colon
  \mathbf{B} \to \mathbb{R}_{\ge 0}$, and we want to show that
  \cref{eq:wmccondition} is satisfied. Let $k_i, k_j \in L$ be such that $k_i
  \in \{ l_i, \neg l_i \}$, $k_j \in \{ l_j, \neg l_j \}$, and $i \ne j$ for
  some $i, j \in [n]$. We then want to show that
  \begin{equation} \label{eq:to_prove3}
    m(k_i \land k_j) = m(k_i)m(k_j)
  \end{equation}
  which is equivalent to
  \begin{equation} \label{eq:to_prove4}
    m(k_i \land k_j) \cdot m(\neg k_i \land \neg k_j) = m(k_i \land \neg k_j)
    \cdot m(\neg k_i \land k_j)
  \end{equation}
  by \cref{lemma:before_theorem}. Then
  \begin{align*}
    \WMC(k_i \land k_j) &= \sum_{\text{atoms } a \le k_i \land k_j} \WMC(a) = \sum_{\text{atoms } a \le k_i \land k_j} \prod_{m \in [n]} w(a_m) \\
                        &= \sum_{\text{atoms } a \le k_i \land k_j} w(a_i)w(a_j) \prod_{m \in [n] \setminus \{ i, j \}} w(a_m) = \sum_{\text{atoms } a \le k_i \land k_j} w(k_i)w(k_j) \prod_{m \in [n] \setminus \{ i, j \}} w(a_m) \\
    &= w(k_i)w(k_j) \sum_{\text{atoms } a \le k_i \land k_j} \prod_{m \in [n] \setminus \{ i, j \}} w(a_m) = w(k_i)w(k_j)C,
  \end{align*}
  where $C$ denotes the part of $\WMC(k_i \land k_j)$ that will be the same for
  $\WMC(\neg k_i \land k_j)$, $\WMC(k_i \land \neg k_j)$, and $\WMC(\neg k_i
  \land \neg k_j)$ as well. But then \cref{eq:to_prove4} becomes
  \[
    w(k_i)w(k_j)w(\neg k_i)w(\neg k_j)C^2 = w(k_i)w(\neg k_j)w(\neg k_i)w(k_j)C^2
  \]
  which is trivially true.
\end{proof}

>>>>

\begin{itemize}
\item The Main Narrative
  \begin{enumerate}
  \item When weights are defined on literals, the measure on the free BA is
    fully independent.
  \item This means that the BA itself must be larger (i.e., have additional
    `meaningless' literals) to turn any probability distribution into an
    independent one.
  \item We show how we can define conditional weights on literals, allowing us
    to encode any probability distribution into a Boolean algebra that's not
    necessarily independent and thus can be smaller.
  \item We demonstrate a specific example of this by presenting a new way to
    encode Bayesian networks into instances of WMC and adapting a WMC algorithm
    (ADDMC) to run on the new format.
  \item We show that this results in significantly faster inference.
  \item We show that our encoding results in asymptotically fewer literals and
    fewer ADDs, and thus a simpler problem.
  \item (Maybe) we experimentally demonstrate a phase transition based on the
    number of variables per ADD.
  \end{enumerate}
\item Potential criticism may be that this is a step backwards and doesn't allow
  us to use SAT-based techniques for probabilistic inference. However, they can
  still be used for the 'theory+query' part.
  \begin{itemize}
  \item Zero-probability weights and one-probability weights can be interpreted
    as logical clauses. This doesn't affect ADDMC but could be useful for other
    solvers.
  \end{itemize}
\item[F] What are the main claims, what are the main takeaways, intuitive [???]
  of theorems to follow. To do this, we appeal to algebraic constructions to
  define the main concepts for introducing measures on Boolean algebras.
\item
  Algorithms\footnote{\url{http://beyondnp.org/pages/solvers/model-counters-exact/}}
  \begin{itemize}
  \item ADDMC \cite{DBLP:conf/aaai/DudekPV20} (rediscovered the
    multiplicativity of BAs in different words) (with optimal settings)
  \item Cachet \cite{DBLP:conf/sat/SangBBKP04}
  \item c2d \cite{DBLP:conf/ecai/Darwiche04}
  \item d4 \cite{DBLP:conf/ijcai/LagniezM17} (closed source, boo!)
  \item miniC2D  \cite{DBLP:conf/ijcai/OztokD15}
  \end{itemize}
\item Notable previous/related work
  \begin{itemize}
  \item Hailperin's approach to probability logic
    \cite{DBLP:journals/ndjfl/Hailperin84}
  \item Nilsson's (somewhat successful) probabilistic logic
    \cite{DBLP:journals/ai/Nilsson86,DBLP:journals/ai/Nilsson93}
  \item Logical induction: a big paper with a good overview of previous attempts
    to assign probabilities to logical sentences in a sensible way
    \cite{DBLP:journals/eccc/GarrabrantBCST16}
  \item Measures on Boolean algebras
    \begin{itemize}
    \item On possibility and probability measures in finite Boolean algebras
      \cite{DBLP:journals/soco/CastineiraCT02}
    \item Representation of conditional probability measures
      \cite{krauss1968representation}
    \end{itemize}
  \end{itemize}
\item Intuitively, a measure is just like a probability, except it's in
  $\mathbb{R}_{\ge 0}$ instead of $[0, 1]$.
\end{itemize}

@@

\begin{itemize}
\item Make up my mind about $a, b$ vs. $x, y$ and stick to it (maybe $x, y$?).
\item Terminology: `with generating set $S$' $\to$ `over $S$'.
\item Notation: if $L$ denotes literals, then it doesn't denote a generating
  set. Literals should be $U$.
\item Describe the parallel between BAs and powersets.
\item Describe \cref{tbl:notation_example} as an example. We will use
  set-theoretic notation for $2^U$ and Boolean-algebraic notation for $2^{2^U}$
  (except for atoms).
\end{itemize}

\begin{table}
  \caption{(in the same order)}
  \label{tbl:notation_example}
  \centering
  \begin{tabular}{lcc}
    \toprule
    & Set-theoretic notation & Boolean-algebraic notation \\
    \midrule
    Atoms (elements of $U$) & $a, b$ & $a, b$ \\
    \rowcolor{gray!10} Models (elements of $2^U$) & $\emptyset, \{a\}, \{b\}, \{a, b\}$ & $\neg a \land \neg b, a \land \neg b, \neg a \land b, a \land b$ \\
    & $\{ \emptyset, \{a\}, \{b\}, \{a, b\} \}$ & $\top$ \\
    & $\{ \emptyset, \{a\}, \{b\} \}, \{ \emptyset, \{a\}, \{a, b\} \}$ & $\neg a \lor \neg b, a \to b$ \\
    & $\{ \emptyset, \{b\}, \{a, b\} \}, \{ \{a\}, \{b\}, \{a, b\} \}$ & $b \to a, a \lor b$ \\
    & $\{\emptyset, \{a\}\}, \{\emptyset, \{b\}\}, \{\emptyset, \{a, b\}\}$ & $\neg b, \neg a, a \leftrightarrow b$ \\
    & $\{\{a\}, \{b\}\}, \{\{a\}, \{a, b\}\}, \{\{b\}, \{a, b\}\}$ & $(a \land \neg b) \lor (b \land \neg a), a, b$ \\
    & $\{\emptyset\}, \{\{a\}\}, \{\{b\}\}, \{\{a, b\}\}$ & $\neg a \land \neg b, a \land \neg b, \neg a \land b, a \land b$ \\
    \multirow{-7}{*}{Formulas (elements of $2^{2^U}$)} & $\emptyset$ & $\bot$ \\
    \bottomrule
  \end{tabular}
\end{table}

Consider the Boolean algebra $(2^{2^U}, \land, \lor, \neg, \bot, \top)$. A
partial order $\le$ on $2^{2^U}$ is defined by $a \le b$ if $a = b \land a$ (or,
equivalently, $a \lor b = b$) for all $a, b \in 2^{2^U}$.

\begin{definition}[\cite{DBLP:books/daglib/0090259,levasseur2012applied}]
  An element $a \ne 0$ of $2^{2^U}$ is an \emph{atom} if, for all $x \in
  2^{2^U}$, either $x \land a = a$ or $x \land a = 0$. Equivalently, $a \ne 0$
  is an atom if there is no $x \in \mathbf{B}$ such that $0 < x < a$.
  Equivalently, $a \ne 0$ is an atom if $|a| = 1$, i.e., $a = \{ u \}$ for some
  $u \in 2^U$ (this is how we will denote atoms in the remainder of the paper).
  Note that atoms in Boolean algebras correspond to models.
\end{definition}

@@

@@

\begin{itemize}
\item[F2] Explain with examples: models $=$ elements [atoms] of algebra.
\item Be careful about mentioning ideals and quotients.
\end{itemize}

\begin{definition} \label{def:wmc}
  Let $U$ be an arbitrary set. Then
  \begin{itemize}
  \item a \emph{measure} is a function $M\colon 2^{2^U} \to \mathbb{R}_{\ge 0}$,
  \item a \emph{weight function} is a function $W\colon 2^U \to \mathbb{R}_{\ge
      0}$.
  \item a weight function is \emph{factored} if $W = \prod_{x \in U} W_x$ for
    some functions $W_x\colon 2^{\{x\}} \to \mathbb{R}_{\ge 0}$.
  \item every weight function induces a measure:
    \[
      M_W(x) = \begin{cases}
        0 & \text{if } x = \bot \\
        W(u) & \text{if } x = \{ u \} \\
        \sum_{\{u\} \le x} W(u) & \text{otherwise.}
      \end{cases}
    \]
    The process of calculating the value of $M_W(x)$ for some $x \in 2^{2^U}$
    with a given definition of $W$ is known as \emph{weighted model counting}.
  \item A measure $M\colon 2^{2^U} \to \mathbb{R}_{\ge 0}$ is
    \emph{factorable} if there exists a factored weight function $W\colon 2^U
    \to \mathbb{R}_{\ge 0}$ that induces $M$.
  \end{itemize}
\end{definition}

\begin{example} \label{example:construction}
  Let $\mathcal{L}$ be a propositional logic with $p$ and $q$ as its only atoms.
  Then $L = \{ p, q, \neg p, \neg q \}$ is its set of literals. Let $w : L \to
  \mathbb{R}_{\ge 0}$ be the \emph{weight function} defined by
  \begin{align*}
    w(p) = 0.3, \\
    w(\neg p) = 0.7, \\
    w(q) = 0.2, \\
    w(\neg q) = 0.8.
  \end{align*}
  Let $\Delta$ be a theory in $\mathcal{L}$ with a sole axiom $p$. Then
  $\Delta$ has two models, i.e., $\{ p, q \}$ and $\{ p, \neg q \}$. The
  \emph{weighted model count} (WMC) \cite{DBLP:journals/ai/ChaviraD08} of $\Delta$ is
  then
  \[
    \sum_{\omega \models \Delta} \prod_{\omega \models l} w(l) =
    w(p)w(q) + w(p)w(\neg q) = 0.3.
  \]

@@

@@

\section{Limitations of Factorable Measures}

@@

\begin{lemma} \label{lemma:before_theorem}
  For any measure $M\colon 2^{2^U} \to \mathbb{R}_{\ge 0}$ and elements $a, b
  \in 2^{2^U}$,
  \begin{equation} \label{eq:to_prove}
    M(a \land b) = M(a)M(b)
  \end{equation}
  if and only if
  \begin{equation} \label{eq:to_prove2}
    M(a \land b) \cdot M(\neg a \land \neg b) = M(a \land \neg b)
    \cdot M(\neg a \land b).
  \end{equation}
\end{lemma}
\begin{proof}
  First, note that $a = (a \land b) \lor (a \land \neg b)$ and $(a \land b)
  \land (a \land \neg b) = 0$, so, by properties of a measure,
  \begin{equation} \label{eq:temp}
    M(a) = M(a \land b) + M(a \land \neg b).
  \end{equation}
  Applying \cref{eq:temp} and the equivalent expression for $M(b)$ allows us
  to rewrite \cref{eq:to_prove} as
  \[
    M(a \land b) = [M(a \land b) + M(a \land \neg b)][M(a \land b) + M(\neg a
    \land b)]
  \]
  which is equivalent to
  \begin{equation} \label{eq:temp6}
    M(a \land b)[1 - M(a \land b) - M(a \land \neg b) - M(\neg a \land b)] = M(a
    \land \neg b)M(\neg a \land b).
  \end{equation}
  Since $a \land b$, $a \land \neg b$, $\neg a \land b$, $\neg a \land \neg b$
  are pairwise disjoint and their supremum is $1$,
  \[
    M(a \land b) + M(a \land \neg b) + M(\neg a \land b) + M(\neg a \land \neg
    b) = 1,
  \]
  and this allows us to rewrite \cref{eq:temp6} into \cref{eq:to_prove2}. As all
  transformations are invertible, the two expressions are equivalent.
\end{proof}

\todo[inline]{This theorem needs a special case for zero weights.}
\begin{theorem}
  A measure $M\colon 2^{2^U} \to \mathbb{R}_{\ge 0}$ is factorable if and only
  if
  \begin{equation} \label{eq:wmccondition}
  M(u \land v) = M(u)M(v)
  \end{equation}
  for all distinct $u, v \in U \cup \{ \neg w \mid w \in U \}$ such that $u \ne
  \neg v$.
\end{theorem}
\begin{proof}
  ($\Leftarrow$) For each $x \in U$, let $W_x\colon 2^{\{x\}} \to
  \mathbb{R}_{\ge 0}$ be defined by $W_x(\{ x \}) = M(x)$ and $W_x(\emptyset) =
  M(\neg x)$. Let $M_W$ be the measure induced by
  \[
    W = \prod_{x \in U} W_x.
  \]
  We will show that $M = M_W$. First, note that $M_w(\bot) = 0 = M(\bot)$ by the
  definitions of both $M_w$ and $M$. Second, let
  \begin{equation} \label{eq:def_of_a}
    a = \bigwedge_{u \in U} a_u
  \end{equation}
  be an atom in $2^{2^U}$ such that $a_u \in \{ u, \neg u \}$ for all $u \in U$.
  Then
  \[
    M_W(a) = \prod_{u \in U} W_u(a_u) = \prod_{u \in U} M(a_u) = M
    \left(\bigwedge_{u \in U} a_u \right) = M(a)
  \] % TODO: by what?
  by \cref{def:wmc,eq:wmccondition,eq:def_of_a}. Finally, note that if $M_W$ and
  $M$ agree on all atoms, then they must also agree on all other non-zero
  elements of the Boolean algebra.

  ($\Rightarrow$) For the other direction, we are given a factored weight
  function
  \[
    W = \prod_{x \in U} W_x,
  \]
  and we want to show that its induced measure $M_W$ satisfies
  \cref{eq:wmccondition}. Let $k_u, k_v \in U \cup \{ \neg w \mid w \in U \}$ be
  such that $k_u \in \{ u, \neg u \}$, $k_v \in \{ v, \neg v \}$, and $u \ne v$.
  We then want to show that
  \begin{equation} \label{eq:to_prove3}
    M_W(k_u \land k_v) = M_W(k_u)M_W(k_v)
  \end{equation}
  which is equivalent to
  \begin{equation} \label{eq:to_prove4}
    M_W(k_u \land k_v) \cdot M_W(\neg k_u \land \neg k_v) = M_W(k_u \land \neg k_v) \cdot M_w(\neg k_u \land k_v)
  \end{equation}
  by \cref{lemma:before_theorem}. Then
  \begin{align*}
    M_W(k_u \land k_v) &= \sum_{\{a\} \le k_u \land k_v} W(a) = \sum_{\{a\} \le k_u \land k_v} \prod_{x \in U} W_x(a) \\
                        &= \sum_{\{a\} \le k_u \land k_v} W_u(a_u)W_v(a_v) \prod_{x \in U \setminus \{ u, v \}} W_x(a) = \sum_{\{a\} \le k_u \land k_v} W_u(k_u)W_v(k_v) \prod_{x \in U \setminus \{ u, v \}} W_x(a) \\
    &= W_u(k_u)W_v(k_v) \sum_{\{a\} \le k_u \land k_v} \prod_{x \in U \setminus \{ u, v \}} W_x(a) = W_u(k_u)W_v(k_v)C,
  \end{align*}
  where $C$ denotes the part of $M_W(k_u \land k_v)$ that will be the same for
  $M_W(\neg k_u \land k_v)$, $M_W(k_u \land \neg k_v)$, and $M_W(\neg k_u \land
  \neg k_v)$ as well. But then \cref{eq:to_prove4} becomes
  \[
    W_u(k_u)W_v(k_v)W_u(\neg k_u)W_v(\neg k_v)C^2 = W_u(k_u)W_v(\neg
    k_v)W_u(\neg k_u)W_v(k_v)C^2
  \]
  which is trivially true.
\end{proof}
